---
title: "EDA, 653 Applied Longitudinal Data Analysis"
author: 
    - "Term Project; Parametric Modeling of the Mean Curves; Covariance Models"
date: '`r Sys.Date()`'
---
```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)

rm(list=ls())

library(tidyverse)
library(reshape2)
library(foreign)
```


```{r}
df<- read.csv("G:/My Drive/BIOS 653/hw03/data_project.csv",
              header = TRUE, 
              sep = ',', 
              stringsAsFactors = TRUE)
colnames(df)[1] <- "Sex" 
df <- df[order(df$ID),]

df$Days[df$Time == 1] <- 0
df$Days[df$Time == 2] <- 93
df$Days[df$Time == 3] <- 96
head(df, 15)

```

```{r}
cor(x = df$Fatmass, y = df$FFM)
plot(x = df$Fatmass, y = df$FFM)
model_fat<-lm(FFM ~ Fatmass, data = df)
summary(model_fat)

covariates <- data.frame(sex=df$Sex, exerc_group=df$Group, time= df$Time, fatmass= df$Fatmass, FFM= df$FFM, musclehsl= df$MuscleHSL)

head(covariates)
library(psych)
covariates_plot <- pairs.panels(covariates)
```

```{r}
MICT_0_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==0)])
MICT_0_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==0)])

HIIT_0_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==1)])
HIIT_0_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==1)])

MICT_93_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Group==0)])
MICT_93_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Group==0)])

HIIT_93_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Group==1)], na.rm = T)
HIIT_93_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Group==1)], na.rm=T)

MICT_96_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==0)])
MICT_96_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==0)])

HIIT_96_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==1)], na.rm = T)
HIIT_96_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==1)], na.rm=T)

MICT_0<-cbind(MICT_0_avg, MICT_0_sd)
MICT_93<-cbind(MICT_93_avg, MICT_93_sd)
MICT_96 <-cbind(MICT_96_avg, MICT_96_sd)

HIIT_0 <- cbind(HIIT_0_avg, HIIT_0_sd)
HIIT_93 <- cbind(HIIT_93_avg, HIIT_93_sd)
HIIT_96 <- cbind(HIIT_96_avg, HIIT_96_sd)

MICT<-rbind(MICT_0, MICT_93, MICT_96)
HIIT<-rbind(HIIT_0, HIIT_93, HIIT_96)

by_trt<-data.frame(MICT=MICT, HIIT=HIIT)
colnames(by_trt)<-c("mean for MICT", "sd for MICT", "mean for HIIT", "sd for HIIT")
row.names(by_trt) <- c("baseline", "93 days", "96 days")
print(by_trt)
```


```{r}
F_0_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Sex==0)])
F_0_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Sex==0)])

M_0_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Sex==1)])
M_0_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Sex==1)])

F_93_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Sex==0)], na.rm = T)
F_93_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Sex==0)], na.rm = T)

M_93_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Sex==1)], na.rm = T)
M_93_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==93 & df$Sex==1)], na.rm=T)

F_96_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Sex==0)])
F_96_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Sex==0)])

M_96_avg <- mean(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Sex==1)], na.rm = T)
M_96_sd <- sd(df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Sex==1)], na.rm=T)

F_0<-cbind(F_0_avg, F_0_sd)
F_93<-cbind(F_93_avg, F_93_sd)
F_96 <-cbind(F_96_avg, F_96_sd)

M_0 <- cbind(M_0_avg, M_0_sd)
M_93 <- cbind(M_93_avg, M_93_sd)
M_96 <- cbind(M_96_avg, M_96_sd)

F<-rbind(F_0, F_93, F_96)
M<-rbind(M_0, M_93, M_96)

by_trt<-data.frame(F=F, M=M)
colnames(by_trt)<-c("mean for F", "sd for F", "mean for M", "sd for M")
row.names(by_trt) <- c("baseline", "93 days", "96 days")
print(by_trt)
```


```{r}

par(mfrow=c(3,3))
hist(x = df$GIRperkgFFMperinsulin[which(df$Days==0)], 
     main = "Histogram for Glucose \nInsulin Resistance at Baseline", 
     xlab= "GIR per kg FFM per insulin")

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==0)], 
     xlab = "GIR per kg FFM per insulin", 
     main = "Histogram for Glucose \nInsulin Resistance at Baseline for MICT")

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==0 & df$Group==1)], 
     xlab = "GIR per kg FFM per insulin", 
     main = "Histogram for Glucose \nInsulin Resistance at Baseline for HITT", 
     ylim = c(0, 6))

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==93)], 
     main = "Histogram for Glucose \nInsulin Resistance at Visit 2", 
     xlab= "GIR per kg FFM per insulin")

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==96)], 
     main = "Histogram for Glucose \nInsulin Resistance at Visit 3", 
     xlab= "GIR per kg FFM per insulin")

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==0)], 
     main="Histogram for Glucose \nInsulin Resistance at Visit 3 for MICT", 
     xlab= "GIR per kg FFM per insulin")

hist(x = df$GIRperkgFFMperinsulin[which(df$Days==96 & df$Group==1)], 
     main="Histogram for Glucose \nInsulin Resistance at Visit 3 for HIIT",
     xlab= "GIR per kg FFM per insulin")
```

```{r}
rm(Visit0)
rm(Visit1)
rm(Visit2)
Visit1<-df$GIRperkgFFMperinsulin[which(df$Days==0)]
Visit2<-df$GIRperkgFFMperinsulin[which(df$Days==93)]
Visit3<-df$GIRperkgFFMperinsulin[which(df$Days==96)]
visits<-data.frame(Visit1, Visit2, Visit3)
pairs.panels(visits, main="Correlations Between Repeated Insulin Resistance Measurements")
```

The response is: GIRperkgFFMperinsulin

The variables names are:

* GIRperkgFFMperinsulin
* ID
* Sex
* Group (0 for moderate intensity, 1 for high intensity)
* Days
* 
The covariates are:

* Group
* Sex
* BMI
* 

4. For continous outcomes, use spaghetti plots to visualize 20 individuals' responses over time; For non-continuous outcomes, plot 10 individuals' responses over time in 10 separate panels.

```{r}
df.spaghetti <- na.omit(df)
df.spaghetti <- within(df.spaghetti, {
  ID <- factor(ID) 
  Group <- factor(Group, levels=0:1, labels = c("MICT", "HIIT"))})
library(ggplot2)


spaghetti <- ggplot(data = df.spaghetti, aes(x = Time, y = GIRperkgFFMperinsulin, group=ID))

spaghetti + geom_line() + stat_smooth(aes(group=1), method = "gam") + facet_grid(.~ df.spaghetti$Group) + xlab("Time") + ylab("Glucose Infusion Rate per kg Fat Free Mass per insulin")

```




