---
title: "653 Applied Longitudinal Data Analysis"
author: 
    - "Term Project"
date: '`r Sys.Date()`'
---
```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)

rm(list=ls())

library(tidyverse)
library(reshape2)
library(foreign)
```

I had to alter the original file to remove space for Muscle glycogen
```{r}
df1<- read.csv("G:/My Drive/BIOS 653/project/data_project.csv",
              header = TRUE, 
              sep = ',', 
              stringsAsFactors = TRUE)
colnames(df1)[1] <- "Sex" 
df1 <- df1[order(df1$ID),]

# Create a variable called Days, represent day of measurements
df1$Days[df1$Time == 1] <- 0
df1$Days[df1$Time == 2] <- 93
df1$Days[df1$Time == 3] <- 96
head(df1, 15)

```

This was in the proposal, just shows that we know there are correlations between repeated measures and some insights on a potential correlation sturcture.

```{r}
Visit1<-df1$GIRperkgFFMperinsulin[which(df1$Days==0)]
Visit2<-df1$GIRperkgFFMperinsulin[which(df1$Days==93)]
Visit3<-df1$GIRperkgFFMperinsulin[which(df1$Days==96)]
visits<-data.frame(Visit1, Visit2, Visit3)
pairs.panels(visits, main="Correlations Between Repeated Insulin Resistance Measurements")
```


Explore relationships between covariates and outcome

* Check for multicollinearity between FFM and Fatmass - looks fine to include in model
* Appears that FFM and Sex are correlated - should potentially drop one (probably FFM)
* Should plot GIRperkgFFMperinsulin vs Fatmass, FFM, muscleglyc, and COVIX by themselves to get a clearer picture

```{r}
cor(x = df1$Fatmass, y = df1$FFM)
plot(x = df1$Fatmass, y = df1$FFM)
model_fat<-lm(FFM ~ Fatmass, data = df1)
```


```{r}
data <- data.frame(ID=df1$ID, Sex=df1$Sex, Group=df1$Group, Days= df1$Days, Fatmass= df1$Fatmass, FFM= df1$FFM, Muscle.glycogen= df1$'Muscle.glycogen', COXIV = df1$COXIV, GIRperkgFFMperinsulin = df1$GIRperkgFFMperinsulin)

head(data)
library(psych)
covariates_plot <- pairs.panels(data[,-c(1)])
```



```{r, message=F}
#install.packages("lme4")
#install.packages("arm")

library(lme4)
library(arm)
```

Data manipulation; could potentially look into imputing missing data.

```{r}
str(data)
data$ID <- as.factor(data$ID) #ID should be treated as 'categorical'
data <- data[-c(94:96),] # these rows are empty
summary(data) 
```

Export the data to use in SAS or other programs.

```{r}
filepath <- "C:/Users/janes/Downloads/data_biostat653_project.csv"
write.table(x = data, file = filepath)
```

This is a model for insulin sensitivity with Days as fixed effects, and intercept for each ID as random effect.

```{r}
mod0<-lmer(GIRperkgFFMperinsulin ~ Days + (1|ID), data)
summary(mod0)
display(mod0)
confint(mod0)
```

This is a model for insulin sensitivity with Sex, Group*Days, Fatmass, FFM, and Muscle glycogen as fixed effects, with a random intercept for each individual and random slope for time. Note there was an issue with convergence.

```{r}
model.lmm<-lmer(GIRperkgFFMperinsulin~Sex+Group*Days+Fatmass+FFM+Muscle.glycogen + (Days|ID), data = data)
summary(model.lmm)
display(model.lmm)
confint(model.lmm)
```

This is a model for insulin sensitivity with Sex, Group*Days, Fatmass, FFM, and Muscle glycogen as fixed effects, with a random intercept for each individual.

```{r}
model2.lmm<-lmer(GIRperkgFFMperinsulin~Sex+Group*Days+Fatmass+FFM+Muscle.glycogen + (1|ID), data = data)
summary(model2.lmm)
display(model2.lmm)
confint(model2.lmm)
```

This is a model for insulin sensitvity with Sex, Group*Days, Fatmass, FFM, Muscle.glycogen, and Sex*Fatmass as fixed effects, with a random intercept for each individual.

```{r}
model3.lmm<-lmer(GIRperkgFFMperinsulin~Sex+Group*Days+Fatmass+FFM+Muscle.glycogen +Sex*Fatmass+ (1|ID), data = data)
summary(model3.lmm)
display(model3.lmm)
confint(model3.lmm)
```

Model fit is not significantly different for these models; smallest AIC is model2.lmm

```{r}
anova(model.lmm, model2.lmm, model3.lmm)
```


Fitted vs. residual plots
Suggest a little issue with more variability at larger fitted values

```{r}
plot(model2.lmm, type = c("p", "smooth"))
```

Quantile-Quantile plots
Residuals are not quite normal.

```{r}
qqmath(model2.lmm, id=0.05)
```

Playing with more models

```{r}
mod1<-lmer(GIRperkgFFMperinsulin~ Sex + Group*Days+Fatmass+Muscle.glycogen + (1|ID), data = data)

display(mod1)
confint(mod1)
anova(model2.lmm, mod1)
```

```{r}
p <- ggplot(data = df, aes(x = COXIV, y =GIRperkgFFMperinsulin, group = ID))
p + geom_line()
```


Some casual Random Forest Feature Selection
```{r}
library(Boruta)
boruta_out<-Boruta(GIRperkgFFMperinsulin~., data=na.omit(df1), doTrace=2)
names(boruta_out$finalDecision[boruta_out$finalDecision %in% c("Confirmed", "Tentative")])
plot(boruta_out, cex.axis=.7, las=2, xlab="", main="Variable Importance")
```
